Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-03-25 demystify.tex
updates

\section{Why did I get rejected?}\label{sec:why-rejected}
Many possible reasons!  Other than not having a strong profile (e.g., research potentials, GPA, LoRs, SOP, poor interviews), here are some that you might not expect.

\paragraph{Not a good fit}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you might not be admitted.
Similarly, if there is no faculty who is willing to advise you (e.g., they are on sabbatical, personal leave, do not have sufficient fundings, or already have too many students already), you will not be admitted.  This is actually good for you, as you don't want to be in a program where no one can advise you.

@@

\paragraph{You aim too high}  You might have applied to schools that are too high for your profile. While being ambitious is good, you should also be realistic and apply to a range of schools, including ``safety'' ones.  I have seen many students who have excellent profiles but only applied to top 10 or even top 20 schools and got rejected from all of them.  Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply

\paragraph{Overqualified or Lack Interest}  This might be paradoxical, however if the adcom believes that you are likely to get admitted and go to a better program (i.e., they are just your ``safety'' schools), they might not be so enthusiastic to admit you and want to save the spot for someone else. Note though it does happen, for most good schools this is rare and you should not assume this is the case.

@@

\paragraph{Red flags in Your Background} Plagiarism, cheating, or other academic dishonesty are \red{red} flags. Another one is if you have a history of jumping from one program to another (e.g., you have been in multiple different programs in short time). If you think you have red flags, you should address them in your SOP.
In general, these things are rare, but they do happen and cause concern to the adcom.

>>>>

\section{Why did I get rejected?}\label{sec:why-rejected}
Many possible reasons!  Other than not having a strong profile (e.g., research potentials, GPA, LoRs, SOP, poor interviews), here are some other reasons.

\paragraph{You aim too high}  This is perhaps the most common one: you have applied to schools that are too high for your profile. While being ambitious is good, you should understand how PhD admission works (e.g., by reading this handbook! and realize things such as having a good GPA doesn't mean much) and be realistic!

You should apply to a range of schools, including ``safety'' ones.   Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply.

\paragraph{Not a good fit}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you might not be admitted.
Similarly, if there is no faculty who is willing to advise you (e.g., they are on sabbatical, personal leave, do not have sufficient funding, or already have too many students already), you will not be admitted.  This is actually good for you, as you don't want to be in a program where no one can advise you.

@@

\paragraph{Overqualified or Lack Interest}  This might be paradoxical and is the opposite of aiming too high. However if the adcom believes that you are likely to get admitted and go to a better program (i.e., they are just your ``safety'' schools), they might not be so enthusiastic to admit you and want to save the spot for someone else. Note though it does happen, for most good schools this is rare and you should not assume this is the case.

@@

\paragraph{Red flags in Your Background} Plagiarism, cheating, or other academic dishonesty are \red{red} flags. Another one is if you have a history of jumping from one program to another (e.g., you have been in multiple different programs in short time). How would the reviewers know?  They might have contacts in other programs and heard about you, your LoRs might mention it.

If you think you have red flags, you should address them in your SOP or ask your letter writers to do so. 
In general, these things are rare, but they do happen and cause concern to the adcom.
